A Switching Planner for Combined Task and Observation Planning

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/discussion.tex
New section

There have been a number of developments recently towards planning
under uncertainty using systems that were intended for sequential
planning in deterministic problems.  Notably,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition.
In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
plans and corresponding execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust sequential plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.
%%
Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
---i.e., sequential planning in unobservable worlds--- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. Most
recently this research thread has been extended to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.

>>>>

A theoretical criticism of switching continual planning, concerns
interleaved sequential and decision-theoretic sessions failing to make
any progress towards the objective.

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
I have to go home to cook dinner for Katrina. Still have to write the section about DTP from serial abstraction..
 - Charles
Might still have some algebra wrong, but I think I did some repairing of the reward sketch.
 - Charles
Indexing error.
 - Charles
New section

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(:init)$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$
$s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
for each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\actions(p_i;(T_i))$ can be executed if no
$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
is in the root conjunct, or it occurs in $T_k$ for some executed
$\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

\[
V^* = \max_{p_1, .., p_n} \prod_{i=1..N} p_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]

@@

\[
V^* = \arg\min_{V_i} \Expect \bigg{[} 
\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate_0
\bigg{]} - V_i.
\]

In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of the infamous degenerate
hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy.

%% Finally, we come to the issue of valuing sequential plans, and
%% therefore giving the criterion that the sequential planner seeks to
%% optimise. Where $p_i$ is the probability that the $i^{th}$ action
%% execution is not a transition to $s^\bot$, we value a trace according
%% to:

%% \begin{equation}\label{eq:tracevalue}
%% V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
%% \end{equation}

%% \noindent In goal directed problems, where the only non-zero reward is
%% received at the first transition to a goal state,
%% Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
%% infamous degenerate hindsight system
%% \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
%% gradient-based reinforcement learning systems, such as William's
%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
%% under reasonable assumptions the optimal trace yields a maximal
%% gradient step from a uniform soft-max policy.

@@

The problem posed to the decision-theoretic system is an abstraction
of the problem at hand with a distinct utility model that focuses the
agents activities on assumption verification. The abstraction is
induced according to the executing sequential plan. Specifically, we
first omit all problem variables that are not mentioned in the linear
trace, and then refine that preliminary model, by progressively adding
more problem variables while the resulting state space remains of a
practicable size. Refinement in our approach is a simple and
inexpensive step, guided by the level at which variable symbols occur
in the description of the initial conditions. The abstract process is
also given a utility model that captures the reward (resp. penalty) of
correctly (resp. falsely) in/validating the assumptions made in the
linear trace. In detail, for each assumption on a state variable, we
can have a reward of $1$ for correctly rejecting it, and a penalty of
$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
a more-or-less continual manner until an assumption is rejected, or
until it returns control to the sequential system because it can see
no benefit from performing further actions in the abstraction.

Addressing briefly the specifics of our mobile robot's
architecture. In our scenario the utility model for the POMDP posed to
decision-theoretic planning is sourced from a motivational component
of our robotic architecture. This proposes a non-zero reward
(resp. penalty) for sequential planning with a true (resp. false)
assumption about a state proposition.

>>>>

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(:init)$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.G., for
$\pp{R2D2}$ $s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\assumptiveS(i)$ can be executed if no $\assumptiveS(j)$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS(k)$.
%%
Executing $\assumptiveS(i)$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

\[
V^* = \max_{p_1, .., p_n} \prod_{i=1..N-1} p_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]

@@

\[
V^* = \arg\min_{V_i} \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
\bigg{]} - V_i.
\]

In goal directed problems, where the only non-zero reward is received
at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
criteria corresponds exactly to that of the infamous degenerate
hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

@@

In the decision-theoretic mode, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ block
from the original problem description.

The focus for decision-theoretic planning is on sensing, rather than
the achievement of utility according to the given objectives. We
therefor give the process in this mode a different reward model, which
reflects the value of performing sensing actions in the context of the
trace proposed for the preceding sequential deliberation. In detail,
all positive rewards from the initial POMDP are removed. Then, for
each $\assumptiveS{i}$ action scheduled in the current sequential
plan, we have a {\em dual} $\assumptiveDT{i}$ so that
$\forall\state\in\states$:

\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\$(T_i) & \pp{otherwise} \\
\end{array}
\]

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly jettisoning a
false assumption. The added dual actions are propositionally trivial
-- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic mode, we include another propositionally trivial
action $\actions.\poss(\action_i)$ with the reward property:

\[
\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
-\$(\poss(\action_i)) & \pp{otherwise} \\
\end{array}
\]

For the management of switching, execution of either a
$\assumptiveDT{i}$ action, or $\actions.\poss(\action_i)$, returns
control to the sequential planner. At that point, if a dual is
executed, then the sequential planner must replan with $\bstate_0$
equal to the current belief. Otherwise, if $\actions.\poss(\action_i)$
is executed, then the current sequential plan is executed until
further sensing is scheduled, or to completion. In our project the
rewards, i.e., function $\$:2^\propositions\to\Re$, for the POMDP
posed to decision-theoretic planning are sourced from a motivational
component of our robotic architecture.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
constructing a tree which only includes terms from the {\em relaxed}
visitation characterised by the {\em assumptive} actions in the
current sequential plan. For example, if $\pp{R2D2}$'s plan is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
(\pp{report-is-in}~\pp{box}~\pp{office})
\end{array}
\]

\noindent Then the {\em relaxed} block will be,

\small
\begin{tabtt}
(\=:init (= (is-in R2D2) kitchen) \+ \\
       (probabilistic \=.8 (= (is-in box) office)  \\
		      \>.2 (= (is-in box) kitchen))) \\
\end{tabtt}
\normalsize

\noindent and therefore, the belief-state $\bstate_0$ is:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
\hline
\end{tabular}
\normalsize

\noindent At this point, decision-theoretic planning would proceed in an
abstraction of the environment that does not contain a $\pp{cup}$,
where the inclusion (exclusion) of POMDP actions is the same as we
defined it for the sequential mode.

In a second step, we iteratively refine the {\em relaxed} block by
adding terms from the original block while the resulting state space
remains of a practicable size. Of course, it is expensive to measure
size exactly, thus in practice we estimate it to be
$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
propositions in the abstraction induced by the proposed tree. Terms
are proposed for addition because they are probabilistic, and because
they were omitted in the {\em relaxed} visitation characterised by
{\em assumptive} actions in the sequential plan. Each iteration, we
add one candidate term with the largest number of parent terms already
in the candidate tree.
